\addcontentsline{toc}{chapter}{Abstract}\vspace{-1cm}
%Border
\begin{tikzpicture}[remember picture, overlay]
	\draw[line width = 4pt] ($(current page.north west) + (0.75in,-0.75in)$) rectangle ($(current page.south east) + (-0.75in,0.75in)$);
\end{tikzpicture}

\begin{center}
	\textbf{\Large Abstract}
\end{center}

Price volatility during monsoons presents major challenges for Indian farmers, especially for weather-sensitive crops like soybean and onion. Traditional price forecasting systems lack integration with real-time weather data and monsoon-specific variables, leading to inaccurate predictions and economic losses. This project proposes a monsoon-driven crop price prediction platform that combines machine learning with satellite imagery, rainfall patterns, and market data to deliver real-time, accurate forecasts.

The system leverages Random Forest, Support Vector Regression, and Gradient Boosting models, optimized through GridSearchCV and cross-validation, to predict prices using NDVI data, rainfall trends, and historical prices across 1000+ markets. Feature engineering and ensemble modeling enhance accuracy and robustness.

Developed using Python (scikit-learn, pandas, numpy), the system features a React.js frontend, Node.js backend, and hybrid PostgreSQL–MongoDB storage. It achieves up to 95\% prediction accuracy with an MAE of 353 INR for key crops like onion and soybean, outperforming conventional methods by 40\%.

Tested across diverse seasons and markets, the platform supports real-time integration, empowering farmers with timely insights for crop selling and planning. Its scalable, web-based architecture makes it accessible to farmers, policymakers, and insurers, offering a proactive solution to monsoon-related agricultural challenges.

\pagebreak
